Case for Support

### 2021-11-17 application/caseForSupport.tex
Updates from Overleaf

    \item[Understanding the brain and behaviour] Systems neuroscientists aim to
        uncover how adaptive behaviour arises from computation in neural
        circuits in different brain regions. Extending the Bonsai ecosystem
        will facilitate new types of experiment that causally connect neural
        circuits with elemental behavioural processes. Long-term, this research
        will help identify missing links between circuit dysfunction and
        behavioural deficits in neurological or mental health disorders, and
        will support the development of new neural circuit biomarkers
        and therapies. The Bonsai ecosystem will also enable  experimental
        testing of emerging theories of brain function and learning, and thus
        contribute to the neuro-inspired ML algorithms that fuel
        intelligent technologies, increasingly important in our society.  The
        enhanced Bonsai platform will also support purely behavioural
        experiments, which track animals' posture, body parts or location in real-time, to investigate impacts of environmental variables
        or captivity on behaviour - crucial for
        animal welfare.

    \item[Improve reproducibility in behavioural and brain sciences] Lack of
        reproducible results slows down scientific progress and incurs personal
        and economic costs to scientists and funders.  In neuroscience,  many
        behavioural and neural animal studies have been difficult to
        reproduce, and experiments are rarely replicated across laboratories.
        By creating a resource that standardises and enables reproducible
        experimental design/control, as well as analysis of neural and
        behavioural data, the Bonsai ecosystem will allow multiple laboratories
        to implement and align experimental and analytical protocols as closely
        as possible.

    \item[Improve efficiency of research] Most neuroscience laboratories
        worldwide develop custom, unique solutions to their experimental
        control and analysis needs. Some laboratories also lack the technical
        know-how to either develop their own software for experimental hardware
        control or write software for state-of-the-art analysis of rich
        neural and behavioural data, increasingly reliant on the latest ML algorithms. The proposed upgrading of the open-source and
        community-empowered Bonsai ecosystem with a toolbox of online and
        offline analytical tools, which can dovetail with closed-loop
        experimental requirements, will (i) allow the deployment of such tools
        into laboratories that lack the necessary expertise, (ii) reduce the need to
        reinvent the same tools in multiple labs (cost/time efficiency) and
        (iii) concomitantly streamline and standardise the experimental data
        processing streams in systems neuroscience.

>>>>

    \item[Understanding the brain and behaviour] Systems neuroscientists aim to
        uncover how adaptive behaviour arises from computation in neural
        circuits across brain regions. Extending the Bonsai ecosystem
        will facilitate new types of experiment that causally connect neural
        circuits with elemental behavioural processes. Long-term, this research
        will help identify missing links between circuit dysfunction and
        behavioural deficits in neurological or mental health disorders, and
        will support the development of new neural circuit biomarkers
        and therapies. Bonsai will also enable experimental
        testing of emerging theories of brain function and learning, thus
        contributing to the neuro-inspired ML algorithms that fuel
        intelligent technologies, increasingly important in our society.  The
        enhanced Bonsai platform will also support purely behavioural
        experiments, which track animals' posture, body parts or location in real-time, to investigate impacts of environmental variables
        or captivity on behaviour - crucial for
        animal welfare.

    \item[Improve reproducibility in behavioural and brain sciences] Lack of
        reproducible results slows scientific progress, increasing costs to scientists and funders.  Many
        behavioural and neural animal studies have been difficult to
        reproduce across laboratories.
        By standardising and enabling reproducible
        experimental design/control, as well as analysis of neural and
        behavioural data, Bonsai will allow multiple laboratories
        to implement and align experimental and analytical protocols as closely
        as possible.

    \item[Improve efficiency of research] Most neuroscience laboratories
        develop custom solutions to their experimental
        control and analysis needs. Some laboratories also lack the technical
        know-how to either develop their own software for experimental hardware
        control or write software for state-of-the-art analysis of rich
        neural and behavioural data, increasingly reliant on the latest ML algorithms. The proposed upgrading of the open-source and
        community-empowered Bonsai ecosystem with a toolbox of online and
        offline analytical tools, which can dovetail with closed-loop
        experimental requirements, will (i) allow the deployment of such tools
        into laboratories that lack the necessary expertise, (ii) reduce the need to
        reinvent the same tools in multiple labs (cost/time efficiency) and
        (iii) concomitantly streamline and standardise the experimental data
        processing streams in systems neuroscience.

### 2021-11-17 application/detailsOfResource.tex
Updates from Overleaf

Spikes from multiple neurons can be detected in recordings from a single
electrode inserted in neural tissue. Spike sorting is the computational step
used to assign each spike to the neuron that generated it. Most spike sorting
algorithms are designed to work offline (i.e., to use a complete recording
session, after the session terminated). Online spike sorting (i.e., the task
of assigning spikes to individual neurons as recordings are being acquired) is a
still unsolved task, especially for recording devices with large number of
electrodes.
%
However, for the type of BCI application proposed here, previous studies have
shown that spike sorting is not beneficial and high performance can be retained
by just detecting spikes, without assigning them to invidivual
neurons~\citep{trautmannEtAl19,todorovaEtAl14}. Thus, we will only detect spikes with a simple zero-corssing method,
and not assign them to individual neurons.  We will, however, integrate into
Bonsai two offline spike sorting methods, Kilosort~\citep[][written in Matlab]{pachitariuEtAl16}, and MountainSort~\citep[][written in
Python]{chungEtAl17}, to provide Bonsai users with spike sorting functionality for
applications that benefit from it.

\paragraph{LFP extraction}
\label{sec:lfpExtractor}

@@

    \item[Gaussian Linear Dynamical
        System]\citep[(GLDS)][]{andersonAndMoore12}: with sufficiently large
        bin sizes, spike counts can be modelled as Gaussian random processes.
        This Gaussianity assumption greatly simplifies the estimation of
        parameters of linear dynamical system (LDS) models, as well as
        inferences about their states. After models' parameters have been
        learned, GLDS can be used online. A unique feature of GLDS is that the
        posterior distribution of states given all observation up to the
        present can be calculated efficiently. This estimate of the posterior
        distribution can be used online for experimental control, as we
        propose in Section~\ref{sec:comparisonOfMultipleMethods}. We will use an R implementation of GLDS
        which allows the use of external
        inputs\footnote{https://github.com/joacorapela/kalmanFilter}.

>>>>

Spikes from multiple neurons can be detected in recordings from a single
electrode inserted in neural tissue. Spike sorting is the computational step
used to assign each spike to the neuron that generated it. Most spike sorting
algorithms are designed to work offline (i.e., to use a complete recording
session, after the session terminated). Online spike sorting (i.e., the task
of assigning spikes to individual neurons as recordings are being acquired) is a
still unsolved task, especially for recording devices with large number of
electrodes.
%
However, for the type of BCI application proposed here, previous studies have
shown that spike sorting is not beneficial and high performance can be retained
by just detecting spikes, without assigning them to invidivual
neurons~\citep{trautmannEtAl19,todorovaEtAl14}. Thus, we will only detect spikes with a simple zero-crossing method,
and not assign them to individual neurons.  We will, however, integrate into
Bonsai two offline spike sorting methods, Kilosort~\citep[][written in Matlab]{pachitariuEtAl16}, and MountainSort~\citep[][written in
Python]{chungEtAl17}, to provide Bonsai users with spike sorting functionality for
applications that benefit from it.

\paragraph{LFP extraction}
\label{sec:lfpExtraction}

@@

    \item[Gaussian Linear Dynamical
        System]\citep[(GLDS)][]{andersonAndMoore12}: with sufficiently large
        bin sizes, spike counts can be modelled as Gaussian random processes.
        This Gaussianity assumption greatly simplifies the estimation of
        parameters of linear dynamical system (LDS) models, as well as
        inferences about their states. After models' parameters have been
        learned, GLDS can be used online. A unique feature of GLDS is that the
        posterior distribution of states given all observation up to the
        present can be calculated efficiently. This estimate of the posterior
        distribution can be used online for experimental control, as we
        propose in Section~\ref{sec:comparisonOfMultipleMethods}. We will use an R implementation of GLDS
        which allows the use of external
        inputs\footnote{https://github.com/joacorapela/ kalmanFilter}.
